cover_letter

### 2021-10-13 cover_letter.Rmd
modify file names for R journal submission

---
title: An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullets
header-includes:
  - \usepackage{amsmath}
  - \usepackage{xcolor}
  - \usepackage{longtable}
  - \usepackage{adjustbox}            
  - \usepackage[font=small,skip=5pt]{caption}
  - \usepackage{subcaption}
  - \usepackage{afterpage}
  - \usepackage{bbm}
author:
  # see ?rjournal_article for more information
  - name: Wangqian W. Ju
    affiliation: Department of Statistics
    address:
    - Center for Statistics and Applications in Forensic Evidence
    - Iowa State University
    - 2438 Osborn Dr
    - Ames, IA 50011
    url: https://github.com/willju-wangqian
    email:  wju@iastate.edu
    orcid: 0000-0002-9977-377X
  - name: Heike Hofmann
    url: https://github.com/heike
    email: hofmann@iastate.edu
    orcid: 0000-0002-9079-593X
    affiliation: Department of Statistics
    address:
    - Center for Statistics and Applications in Forensic Evidence
    - Iowa State University
    - 2438 Osborn Dr
    - Ames, IA 50011
thanks: |
  This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement 70NANB20H019 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, Duke University, University of California Irvine, University of Virginia, West Virginia University, University of Pennsylvania, Swarthmore College and University of Nebraska, Lincoln.    
abstract: >
  In this paper, we introduce the R package cmpsR, an open-source implementation of the Congruent Matching Profile Segments (CMPS) method developed at the National Institute of Standards and Technology (NIST) for objective comparison of striated tool marks. The functionality of the package is showcased by examples of bullet signatures that come with the package. Graphing tools are implemented in the package as well for users to assess and understand the CMPS results. Initial tests were performed on bullet signatures generated from two sets of 3D scans in the Hamby study under the framework suggested by the R package bulletxtrctr. New metrics based on CMPS scores are introduced and compared with existing metrics. A measure called Sum of Squares Ratio is included, and how it can be used for evaluating different scans, metrics, or parameters is showcased with the Hamby study data sets. An open-source implementation of the CMPS algorithm makes the algorithm more accessible, generates reproducible results, and facilitates further studies of the algorithm such as methods comparison.
preamble: |
  % Any extra LaTeX you need in the preamble

bibliography: CMPSpaper-writing.bib
output: rticles::rjournal_article
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

```{=tex}
\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
\newcommand{\wj}[1]{{\textcolor{blue}{#1}}}
\setlength{\parindent}{0pt}
```

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(tidyverse)
library(bulletxtrctr)
library(kableExtra)
library(patchwork)
library(ggpubr)

source("../code/func_collection.R")

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "\\textwidth",
  fig.align = "center",
#  cache=TRUE,
  dpi = 100,
  warning = FALSE
)
```

```{r setup_knitr, echo=FALSE}
## From Josh O'Brien's stackoverflow answer:
## http://stackoverflow.com/questions/11030898/knitr-how-to-align-code-and-plot-side-by-side
## These two settings control text width in codefig vs. usual code blocks
partWidth <- 35
fullWidth <- 60
options(width = fullWidth)

##  (1) CHUNK HOOK FUNCTION
##   First, to set R's textual output width on a per-chunk basis, we
## need to define a hook function which temporarily resets global R's
## option() settings, just for the current chunk
knit_hooks$set(r.opts = local({
    ropts <- NA
    function(before, options, envir) {
        if (before) {
            ropts <<- options(options$r.opts)
        } else {
            options(ropts)
        }
    }
}))

## (2) OUTPUT HOOK FUNCTION

##   Define a custom output hook function. This function processes _all_
## evaluated chunks, but will return the same output as the usual one,
## UNLESS a 'codefig' argument appeared in the chunk's header.  In that
## case, wrap the usual textual output in LaTeX code placing it in a
## narrower adjustbox environment and setting the graphics that it
## produced in another box beside it.

defaultChunkHook <- environment(knit_hooks[["get"]])$defaults$chunk

codefigChunkHook <- function(x, options) {
  main <-  defaultChunkHook(x, options)
  before <-
    "\\vspace{1em}
\\begin{adjustbox}{valign=t}
\\begin{minipage}{.39\\textwidth}\n"
  after <-
    paste("\\vspace{1em}
\\end{minipage}
\\begin{minipage}{.59\\textwidth}",
    paste0("\\includegraphics[width=\\textwidth]{CMPSpaper_writing_files/figure-latex/",
                 options[["label"]], "-1.pdf}
\\end{minipage}
\\end{adjustbox}"),
          sep = "\n")
  ## Was a codefig option supplied in chunk header?
  ## If so, wrap code block and graphical output with needed LaTeX code.
  if (!is.null(options$codefig)) {
    main <- gsub("```","", main)
    main <- gsub("=latex","", main)
    main <- paste0("{\\small ", main, "}")
    return(sprintf("%s\n%s\n%s", before, main, after))
  } else {
    return(main)
  }
}

knit_hooks[["set"]](chunk = codefigChunkHook)

## (3) TEMPLATE
##   codefig=TRUE is just one of several options needed for the
## side-by-side code block and a figure to come out right. Rather
## than typing out each of them in every single chunk header, we
## define a _template_ which bundles them all together. Then we can
## set all of those options simply by typing opts.label="codefig".

opts_template[["set"]](
codefig = list(codefig = TRUE, fig.show = "hide",
               r.opts = list(width = partWidth),
               tidy.opts = list(width.cutoff = partWidth)))
```

## Introduction

In this paper we present an open-source implementation of the algorithm of the Congruent Matching Profile Segments (CMPS) method.
@cmps developed the CMPS method for "objective comparison of striated tool marks" and demonstrated its use in some examples of comparing bullet signature correlations.
Although @cmps conceptually described the CMPS algorithm in their paper, the authors did not release an implementation of their method.
Thus, our effort here is to introduce the CMPS method to the `R` [@R] community and provide an open-source, publicly available implementation of the algorithm to use, review, and improve.
Our implementation is made available as part of the `R` package \pkg{cmpsR} on GitHub at \url{https://github.com/willju-wangqian/cmpsR}.

According to the Uniform Crime Reporting Program of FBI (Federal Bureau of Investigation), "More than 73 percent (73.7) of the homicides for which the FBI received weapons data in 2019 involved the use of firearms" [@fbiucr], and firearm examination could play an important role in all these cases. \hh{XXX the new data for 2020 came out on Sep 27 - homicides are up by close to 30%. We might want to mention that.}
<!-- Firearm examination has  played an important role in  convictions [@fbiucr] in the United States criminal justice system. \hh{Do we have a number or a percentage of cases where firearm examination has played a role? or maybe gun violence? XXX} --> <!-- https://ucr.fbi.gov/crime-in-the-u.s/2019/crime-in-the-u.s.-2019/topic-pages/expanded-homicide --> 
An important task of firearm examination is to answer the question of whether two pieces of evidence come from the same source or whether a piece of evidence matches a sample obtained from a specific firearm.
Here, in particular, we are interested to determine whether two bullets were fired from the same gun barrel.
Assessing the similarity between two bullets is based on a comparison of striation marks acquired during the firing process as bullets are propelled through the barrel.
Current state of art sees firearms examiners make an assessment of similarity based on a visual comparison, generally, using a comparison microscope [@afte].
This practice has been criticized for its lack of objectivity and the associated problem in calculating an error rate [@pcast].

A report published by @nrc states that "[m]uch forensic evidence-including, for example, bite marks and firearm and toolmark identification---is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline." To overcome those criticisms and concerns, researchers have been making an effort to build databases and develop frameworks and algorithms that bring an objective and quantitative assessment into the field [@nistdb, @brundage, @hamby, @Hamby:2019, @song2005, @ChumbleyL_Scott2010VoTM, @aoas, @pmid30444940; @cmps].
Note that this is not an exhaustive list of these efforts.
The notion of bullet signatures is one of the results of these efforts.
Bullet signatures can be extracted from bullet land engraved areas (LEAs), typically one for each land engraved area, and how those bullet signatures are extracted from scans will be discussed in the Background section.
Since bullet signatures can numerically describe the striation marks on the bullet and can be understood by algorithms, they played an important role in bringing objectivity into the field.
In the following sections we will: review the background of bullet signature comparisons, discuss how we followed the idea described by @cmps for the implementation of the CMPS algorithm, propose new metrics based on the CMPS score and a principled evaluation framework for algorithmic results comparison,  and present results of applying our implementation to real data.

## Background

**Hamby Data Set**
<!-- ### Hamby Data Set -->

The datasets we worked with come from the James Hamby Consecutively Rifled Ruger Barrel Study [@brundage; @hamby; @Hamby:2019], in particular, Hamby set 252 and Hamby set 44.
For each Hamby set, a total of 35 bullets is fired from (the same) ten consecutively manufactured Ruger P-85 pistol barrels.
Two bullets are fired from each barrel, making up a set of 20 reference bullets.
An additional 15 bullets are fired from these ten barrels in a fashion unknown to the study participant.
The aim of the Hamby Study was to have firearms examiners identify which barrel each of the 15 questioned bullets was fired from.
The Ruger P-85 barrels are traditionally rifled barrels with six grooves and lands as shown in \autoref{fig:bullet}.
During the firing process, grooves and lands are engraved on a bullet.
Firearms examiners use striation marks on land engraved areas (LEAs) for their visual comparison.
For algorithmic purposes, 3D topographical images of land engraved areas were obtained and stored in x3p format (XML 3-D Surface Profile).
The x3p format provides a standard way of exchanging 2D and 3D profile data.
It conforms to the ISO5436-2 standard [<http://sourceforge.net/p/open-gps/mwiki/X3p/>] adopted by the OpenFMC (Open Forensic Metrology Consortium), a group of firearm forensics researchers who contributes to the establishment of best practices of using metrology in forensic science.
Hamby set 252 were scanned using a NanoFocus lens at 20x magnification with the scan resolution being 1.5625 \textmu m $\times$ 1.5625 \textmu m per pixel.
Hamby set 44 was scanned at the Roy J Carver High-Resolution microscopy lab at Iowa State.
These scans were acquired with a Sensofar Confocal Light Microscope at 20x magnification for a nominal resolution of 0.645 \textmu m $\times$ 0.645 \textmu m per pixel.
Both Hamby set 252 and Hamby set 44 are publicly available on the NIST Ballistics Database Project [@nistdb].

**Extracting signatures from LEA scans**
<!-- ### Extracting signatures from LEA scans -->

The automated framework used in this paper was proposed by @aoas.
In order to obtain bullet signatures from the x3p files of bullet land engravings, we use the `R` packages `x3ptools` [@x3ptools] and `bulletxtrctr` [@bulletxtrctr].
`x3ptools` is a package to read, write, and generally, process x3p files.
The `bulletxtrctr` package implements a pipeline for extracting and comparing signatures from scans of land-engraved areas.

\autoref{fig:process} gives an overview of all of the steps in the process from scan to signatures.
\autoref{fig:process}(a) shows a rendering of a 3d scan of a bullet land engraved area.
The raised portion of the surface on the left and right of the scan are parts of the adjacent groove engraved areas (GEAs), the middle area shows a land-engraved area (LEA) with well expressed striation marks.
The first step of obtaining the bullet signature is to extract a cross-section at a fixed height along the land engraving.

The white horizontal line in \autoref{fig:process}(a) indicates which cross-section was identified by the algorithm to represent the LEA; \autoref{fig:process}(b) shows the corresponding cross-sectional view.
Groove engraved areas are removed from the analysis as indicated by the vertical blue lines in \autoref{fig:process}(c) and \autoref{fig:process}(d)).
A non-parametric LOESS smooth [@loess] is fitted to capture the bullet curvature (\autoref{fig:process}(e)) and, finally, the **bullet signature** (\autoref{fig:process}(f)) is obtained as residuals of the cross-section and the fitted smooth.
Note, that in @cmps bullet signatures are referred to as bullet profiles.
However, to avoid confusion, we distinguish the notion of bullet signatures from bullet profiles.
Bullet profiles are shown in panels (b), (c), and (d) of \autoref{fig:process}, while \autoref{fig:process}(f) shows the corresponding bullet signature.
Identifying the groove engraved areas correctly is fundamental for a correct down-stream analysis of the signatures.
To allow for a human-in-the middle inspection and intervention we have provided an interactive web-application, implemented as `R` Shiny App [@shiny], `bulletinspectR` to identify and correct those errors.
An example of the extraction process with corresponding code and parameter settings can be found on Github at \url{https://github.com/willju-wangqian/CMPSpaper/tree/main/reproducible}. Note that the process of extracting signatures might be different from the one used in @cmps because no code or parameter settings are made available publicly.

```{r bullet, echo=FALSE, out.width=".8\\textwidth", fig.cap="Photo of a traditionally rifled gun barrel (left) and a fired bullet (right)."}
knitr::include_graphics("img/barrel_bullet_ps.png", dpi = 100)
```

```{r process, echo=FALSE, out.width=".9\\textwidth", fig.cap="A framework of obtaining a bullet signature. (a) front view of a scanned land engraved area (LEA). The selected crosscut location is indicated by the white line. (b) view of the cross-section of the land engraved area at the white line in (a). (c) the crosscut data plotted in 2D; blue vertical lines indicate the position of left and right grooves. (d) the crosscut data after chopping the left and right grooves. (e) the fitted curvature using LOESS. (f) after removing the curvature from the crosscut data, the bullet signature is obtained"}
knitr::include_graphics("img/figure1_v2.PNG", dpi = 100)
```

**Conceptual idea of CMPS**
<!-- ### Conceptual idea of CMPS -->

Most algorithms for comparing striation marks are based on the digitized signatures and produce a similarity score [@song2005, @ChumbleyL_Scott2010VoTM, @aoas, @pmid30444940].
The congruent matching profile segments (CMPS) algorithm, developed by @cmps for "objective comparison of striated tool marks", is one such algorithm.
The algorithm's main idea is to take a set of consecutive and non-overlapping basis segments from the comparison and for each segment find the "best" registration position on the reference (the other bullet signature) with respect to their cross-correlation values.
From a comparison of these registration positions, a **congruent registration position** is identified, and the number of basis segments taking the congruent registration position is the CMPS score.
Note that researchers in @cmps refers to the origin of basis segments as the reference, but in this paper we refer to it as the comparison.
High CMPS scores are achieved between more similar signatures and are therefore indicative of a same-source pair.
Low scores between pairs of signatures are attributed to different source pairs.
However, a specific threshold of the CMPS score to distinguish between same-source and different-source comparisons is not provided in @cmps, instead the threshold depends on the underlying structure of the data and the choice of parameters. \hh{XXX  In a legal setting this variability is problematic because it allows for situations in which experts could choose parameters based on a wished-for outcome. }
Further research is needed to understand how to determine optimal threshold settings.
The CMPS algorithm can assist firearm examiners with drawing a conclusion about the source of a comparison pair.
Unfortunately, @cmps did not release any code or specific parameter settings for their implementation of the CMPS algorithm.
Thus, in this paper we present an open-source implementation of the CMPS algorithm in the R package \pkg{cmpsR} available from Github \url{https://github.com/willju-wangqian/cmpsR}. This publicly available implementation calculates the CMPS score of a comparison using the following code:

```{r, eval=FALSE}
# install.packages("devtools") 
# devtools::install_github("willju-wangqian/CMPS")

library(cmpsR)
data(bullets)

sig1 <- bullets$sigs[[2]]$sig
sig2 <- bullets$sigs[[9]]$sig
sig3 <- bullets$sigs[[10]]$sig

cmps.result.KM <- extract_feature_cmps(sig1, sig2)
cmps.result.KNM <- extract_feature_cmps(sig1, sig3)
```

In this example, the comparison between `sig1` and `sig2`, two signatures coming from the same source (a known-match comparison), gets a CMPS score of 17; the comparison between `sig1` and `sig3`, two signatures coming from different sources (a known non-match comparison), gets a CMPS score of 1.

We also implemented graphing tools for users to better understand these results as well as the algorithm itself.

The section "Implementation" will go through the algorithm and show how to use the \pkg{cmpsR} package. 
A further example that illustrates the main points is also included.
The section "Results" presents the results of evaluating the \pkg{cmpsR} package using Hamby set 252 and Hamby set 44. The results from Hamby set 252 are used to verify that our implementation is, at least qualitatively, comparable to the algorithm described in @cmps. Results from Hamby 44 show the need for a further investigation of the parameter choices even in the case of bullets fired from the same barrels. \wj{We suggest new CMPS metrics that summarize land-level CMPS scores and a sum of squares ratio that can be used to evaluate algorithmic results.}
<!-- \hh{We suggest a sum of square ratio ...} -->
The last section covers some final discussion and conclusions.

## Implementation

**Algorithm**

Conceptually, the CMPS algorithm consists of three main steps:

1.  **cut the comparison signature into consecutive, non-overlapping and equal-length basis segments:** The command `get_segs(x, len=50)` implements this step: it takes bullet signature `x` in the format of a numeric vector and cuts it into consecutive, non-overlapping and equal-length segments of length `len`, which are referred to as "basis segments".

2.  **identify candidate positions:** For each basis segment a set of candidate registration positions on the comparison signature is identified based on the segment's similarity to the reference signature.
    In a first step, the cross-correlation function of the segment to the reference is calculated, then a number of positions with high correlation values are identified as candidate positions.
    In case multiple segment lengths are considered, the length of each basis segment is expanded (by default it is doubled) and these two steps are repeated.
    Only when candidate positions coincide (or are similar enough), they are considered further.
    \autoref{fig:segplots} and \autoref{fig:seg_all} illustrate these ideas.

    -   **Calculate the cross-correlation curve:** Calculate the cross-correlation curve between a basis segment `x` and the reference signature `y` using the function `get_ccf4(x, y, ...)` as shown in \autoref{fig:segplot2}. The position indicates the lag by which a basis segment is moved with respect to its original placement. A position is considered "good" if it results in a peak in the cross-correlation between the basis segment and the reference.

    -   **Correlation peaks:** Two strategies referred to as "multi-peak inspection" and "multi-peak inspection at different segment lengths" in @cmps are used for identifying positions of correlation peaks as candidate positions. The latter is also called "multi-segment lengths strategy". The parameter `npeaks.set` in `extract_feature_cmps(...)` determines which strategy to use and the number of candidate positions:

        -   If `npeaks.set` is an integer vector of length 1, for example `npeaks.set = 5`, the positions of the top five peaks in the cross-correlation curve are identified as candidate positions for registration.
        -   If `npeaks.set` is an integer vector of length more than 1, for example `npeaks.set = c(5, 3, 1)`, the multi-segment lengths strategy will be used: calculate the cross-correlation function between a basis segment and the reference and identify positions of the top five peaks; double the segment length to a specified value, re-calculate the cross-correlation function, and identify three peaks; repeat this process and identify a single peak in the newly computed cross-correlation function. \autoref{fig:seg_all} shows an example of three levels of basis segment 6 and their corresponding cross-correlation curves and identified peaks. Note that in @cmps the segment length is doubled at each level of a basis segment, but in the present implementation users are allowed to choose the segment length at each level.
        -   `get_ccr_peaks(comp, segments, seg_outlength, nseg = 1, npeaks = 5)` computes the cross-correlation curve between a basis or increased segment and the reference signature and finds peaks in the cross-correlation curve. The number of peaks detected is equal to `npeaks`, which is an integer. `segments`, `seg_outlength`, and `nseg` determines the segment in the cross-correlation computation, and `comp` gives the reference signature. If the multi-segment lengths strategy is used, then `get_ccr_peaks(...)` is called in a `lapply()` for each level of the basis segment. The resulting list is called `ccr.list`.

    -   **multi-segment lengths strategy:** with the multi-segment lengths strategy being used, a position is identified as a candidate position for registration and is called as a "consistent correlation peak" if it results in a top peak in the cross-correlation curve with a tolerance zone determined by `Tx` in all segment levels. Note that in @cmps, a segment at its largest scale (highest level) always identifies one peak, but we do not have this requirement in our implementation.

        -   the function `get_seg_scale(segments, nseg, out_length)` is used to obtain the (potentially increased) version of a basis segment. `segments`, which is a list containing all basis segments generated by the function `get_segs(...)` in step 1, and `nseg` are used to determine the basis segment to be increased. `out_length` specifies the length of the output segment.
        -   `get_ccp(ccr.list, Tx = 25)` tries to identify the "consistent correlation peak". `ccr.list` is the result of `lapply()` and `get_ccr_peaks(...)`, and `Tx` determines the size of a tolerance zone used in identifying the consistent correlation peak. `get_ccp(...)` returns `NULL` if there is no consistent correlation peak.

3.  **determine the congruent registration position:** A candidate position "receives" `votes` from basis segments that identify it or a close position within a tolerance zone of `Tx` as a candidate position in step 2.
    Votes for all candidate positions are tallied, and the position with the highest number of votes gets chosen as the *congruent registration position*, indicating that most of the basis segments find their highly similar counterpart in the reference signature in terms of correlation at this registration position.
    In case of ties the middle position is taken as the congruent registration position.
    Basis segments with a congruent registration position are called "congruent matching profile segments" (CMPS).
    The total number of CMPS is the CMPS score of the comparison.
    `get_CMPS(input.ccp, Tx = 25)` is the function that tallies the votes and determines the congruent registration position and congruent matching profile segments (CMPS).

Note that there are several parameters in the CMPS algorithm that are left to the users to decide, such as the length of basis segments `seg_length` in step 1, the number of peaks `npeaks.set` identified on each level in step 2, and the length of the tolerance zone `Tx` in both step 2 and 3.
In our implementation of the CMPS algorithm we used the parameters given in the original CMPS paper [@cmps] as the default values for these parameters.
However, the authors state that no cross-validation has been done - there might be also issues with respect to resolution of the scans.
Further research is needed, until then users are advised to think of default values as starting values and consider alternatives.
The main function that combines all steps in the CMPS algorithm described above is called `extract_feature_cmps(...)`.
Here we present it with its default parameters.

```{r, eval=FALSE}
extract_feature_cmps(
  x,
  y,
  seg_length = 50,
  Tx = 25,
  npeaks.set = c(5, 3, 1),
  include = NULL,
  outlength = NULL
)
```

The function `extract_feature_cmps` allows for the following input from users besides the previously discussed parameters `seg_length`, `npeaks.set`, and `Tx`:

-   `x` and `y` are two signatures: `x` serves as the comparison signature (which will be divided into basis segments) and `y` is the reference signature;

-   `include` determines the format of the function result.
    Besides the `CMPS.score`, other aspects of the comparison help in understanding how the `CMPS.score` is computed.
    By default `include` is set to `NULL` and only the `CMPS.score` is returned; further results are included when `include` is (an abbreviation of) one of or a vector of the following strings: `"nseg", "congruent.pos", "congruent.seg.idx", "segments", "parameters"`, and `"full_result"`.
    If `include` is specified as `"full_result"` (or its abbreviation), the output includes everything listed below.

    -   `nseg`: the number of basis segments from the comparison signature; this is also the highest possible CMPS score of the comparison;
    -   `congruent.pos`: the congruent registration position;
    -   `congruent.seg.idx`: the indices of all congruent matching profile segments;
    -   `ccp.list`: a list showing identified candidate positions of all basis segments;
    -   `pos.df`: a data frame containing all candidate positions and their respective number of votes;
    -   `segments`: a list containing all basis segments;
    -   `parameters`: a list containing all input arguments of `extract_feature_cmps`;

-   `outlength` specifies the segment length of a basis segment at each level under the multi-segment lengths strategy.
    By default `outlength` is set to `NULL`, indicating that a basis segment should double its segment length at the next level and conforming to the description in @cmps .

<!--For more details, please check the documentation of `extract_feature_cmps` or use `?extract_feature_cmps`.-->

<!-- [examples https://journal.r-project.org/archive/2019/RJ-2019-044/RJ-2019-044.pdf] -->

<!-- [https://journal.r-project.org/archive/2019/RJ-2019-002/RJ-2019-002.pdf] -->

In the remainder of the paper we showcase the use of the \pkg{cmpsR} functionality on some examples and present the results of applying it to two datasets.

**Installation**

The \pkg{cmpsR} package is not publicly available from CRAN yet, and its development version is available from Github and can be installed by

```{r, eval=FALSE}
# install.packages("remotes") 
remotes::install_github("willju-wangqian/cmpsR")
```

**An Example**

The \pkg{cmpsR} package contains a simple example to illustrate the basic usage of the package.
The data in this example are twelve bullet signatures obtained from two bullets in Hamby set 252 [@hamby].
The procedure for generating signatures from high-resolution 3D topographic scans of bullet lands used here follows the methodology described in @aoas (as discussed above).
The two bullets under consideration are known to have been fired from the same gun barrel, so for the 36 pairwise land-by-land comparisons, six comparisons are from same-source pairs (known matches) while thirty are from different-source pairs (known non-matches).
To access the example data, we use

```{r, eval=TRUE}
library(cmpsR)
data(bullets)
```

`bullets$sigs` is a list of twelve numeric vectors corresponding to the twelve bullet signatures shown in \autoref{fig:sigs}.
`bullets$source` contains the URLs to the corresponding x3p file containing the topographic scan from the NIST Ballistics Toolmark Research Database [@nistdb].

```{r sigs, echo=FALSE, fig.cap="Signatures of all lands of bullet 1 in the top row, and of bullet 2 in the bottom row. Signatures in the second row are ordered to be in phase with the signatures above, i.e. matching signatures are displayed on top of each other."}
signatures <- bullets %>% unnest(sigs)
signatures <- signatures %>% mutate(
  bulletland = factor(bulletland, levels=c(paste(rep(c(1,2), each=6), c(1:6,2:6,1), sep="-")))
)
signatures %>% ggplot(aes(x = x/1000, y = sig)) + geom_line() + facet_wrap(~bulletland, ncol=6) +
  theme_bw() +
  xlab("Length in mm") +
  ylab("Relative height in micron")
```

The signatures of Land 4 of Bullet 1 and Land 5 of Bullet 2 are stored in objects `sigs2` and `sigs1`, respectively.
This comparison consists of a pair of signatures that are known to be a match -- a KM (known match) comparison.
We compute the CMPS score using two versions of the CMPS algorithm:

```{r, eval=TRUE}
sigs1 <- bullets$sigs[bullets$bulletland == "2-5"][[1]]
sigs2 <- bullets$sigs[bullets$bulletland == "1-4"][[1]]

# compute cmps

# algorithm with multi-peak insepction at three different segment levels
cmps_with_multi_scale <- 
  extract_feature_cmps(sigs1$sig, sigs2$sig, 
                       npeaks.set = c(5,3,1), include = "full_result")

# algorithm with multi-peak inspection at the basis scale only
cmps_without_multi_scale <- 
  extract_feature_cmps(sigs1$sig, sigs2$sig, 
                       npeaks.set = 5, include = "full_result")
```

In the first example, `npeaks.set` is a vector of three integers, i.e. the algorithm uses the multi-segment lengths strategy to create the result object `cmps_with_multi_scale`.
For `cmps_without_multi_scale` each basis segment is linked to the top 5 candidate positions.
We use `include = "full_result"` to capture all results.
In this example, the CMPS is 9 when using multiple segments, and 12 when using a single segment.
As discussed in @cmps, using multi-segment lengths strategy can reduce the number of false positives when identifying candidate positions; however, any score based method is walking the line between false positives and false negatives.
As the number of false positives is reduced the number of false negatives might rise.
More discussion and comparisons between the two versions of the CMPS algorithm will be presented in later sections.
Note, that the multi-segment lengths method is slower because the algorithm is run once for each segment length.

```{r alternative, echo = FALSE}
lands <- unique(bullets$bulletland)

comparisons <- data.frame(expand.grid(land1 = lands[1:6], land2 = lands[7:12]), 
                          stringsAsFactors = FALSE)

comparisons <- comparisons %>% 
  left_join(bullets %>% select(bulletland, sig1=sigs),
            by = c("land1" = "bulletland")) %>%
  left_join(bullets %>% select(bulletland, sig2=sigs),
            by = c("land2" = "bulletland"))

comparisons <- comparisons %>% mutate(
  cmps = purrr::map2(sig1, sig2, .f = function(x, y) {
    extract_feature_cmps(x$sig, y$sig, include = "full")
    })
)

comparisons <- comparisons %>% 
  mutate(
    cmps_score = sapply(comparisons$cmps, function(x) x$CMPS.score),
    cmps_nseg = sapply(comparisons$cmps, function(x) x$nseg),
    cmps_scaled = cmps_score / cmps_nseg
  )

comparisons <- comparisons %>%
  mutate(
    bulletA = gsub("(\\d)-\\d", "\\1", land1),
    landA = gsub("\\d-(\\d)", "\\1", land1),
    bulletB = gsub("(\\d)-\\d", "\\1", land2),
    landB = gsub("\\d-(\\d)", "\\1", land2)
  )

dframe <- comparisons %>% select(-sig1, -sig2)

dframe$samesource <- with(dframe, bullet_to_land_predict(land1=landA, land2=landB, cmps_score, difference=1))
```

**Visualize and Understand CMPS results**

We also implemented graphing tools for visualizing results of the CMPS algorithm.
The goal is to provide users with tools to inspect each of the basis segments and to help them have a better understanding of how the algorithm works.
\autoref{fig:sigplots} shows the plots generated by the first graphing function, `cmps_signature_plot()`, and continues with the example above.
`cmps_signature_plot()` takes the output of `extract_feature_cmps(..., include = "full_result")` and returns a list of 5 elements.
It creates an overall impression of how the comparison signature aligns with the reference signature at the congruent registration position.

-   The first element is a plot called `segment_shift_plot`, shown in \autoref{fig:sigplot}. On this plot the reference signature is drawn as a black line, congruent matching profile segments from the comparison signature are overlaid in red at the congruent registration position.

-   The second plot is called `signature_shift_plot`, shown in \autoref{fig:sigplot2}. This visual presents both the comparison signature and the reference signature. The comparison signature is aligned with the reference signature based on the congruent registration position. Congruent matching profile segments are highlighted by solid red lines.

```{=tex}
\begin{figure}[hbt]
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:sigplot}The black line shows the comparison signature; each red line segment shows one congruent matching profile segment.}
```{r sigplot, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE}
sig.plot <- cmps_signature_plot(
  cmps_with_multi_scale
)
sig.plot$segment_shift_plot
```
\end{subfigure}
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:sigplot2}The black line shows the reference signature; the red line shows the comparison signature. Solid part shows the congruent matching profile segments, and the dashed part shows segments that do not agree with the congruent registration position.}
```{r sigplot2, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE}
sig.plot$signature_shift_plot
```
\end{subfigure}
\caption{\label{fig:sigplots} Two visual outputs of cmps\_signature\_plot()}
\end{figure}
```
-   Other elements of this list are `seg_shift` and `sig_shift`. `sig_shift` gives the congruent registration position, while `seg_shift` is a data frame showing the congruent matching profile segments and their identified candidate position closest to the congruent registration position.

```{r, echo=TRUE, eval=TRUE}
sig.plot$seg_shift
```

While `cmps_signature_plot()` focuses on the signature level, `cmps_segment_plot()` focuses on the segment level.
It provides the "full result" of `extract_feature_cmps()`, but also takes an argument, `seg.idx`, indicating which segment should be inspected.
When checking `sig.plot$seg_shift` we notice that segment number 6 is not one of the congruent matching profile segments.
We can therefore set `seg.idx = 6` in `cmps_segment_plot()` and investigate the reason why this segment disagrees with the congruent registration position.

For each segment scale, we have two plots: `segment_plot` and `scale_ccf_plot`, as shown in \autoref{fig:segplots} for the example of segment number 6:

-   \autoref{fig:segplot} is the `segment_plot` for basis segment 6 at level one (in its original length). We used `npeaks.set = c(5, 3, 1)` in `extract_feature_cmps()` when calculating the CMPS score. Therefore the top five peaks are identified in the cross-correlation curve at level one. Segment 6 is plotted at the positions where these 5 peaks are identified with dashed lines in the `segment_plot`. The solid thick black line shows the segment at its original position (which in this example is very close to the actual registration position).
-   \autoref{fig:segplot2} is the `scale_ccf_plot` of basis segment 6 at level one. It shows the cross-correlation curve computed by the reference signature and the level-one basis segment 6. The five highest peaks are marked by dots on the curve.

```{=tex}
\begin{figure}
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:segplot}segment\_plot for segment 6 at level one. The original position of segment 6 is indicated by the solid black line. Positions where the segment achieves the 5 highest cross-correlations are indicated by the dashed line segments.}
```{r segplot, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE}
seg.plot <- cmps_segment_plot(
  cmps_with_multi_scale, 
  seg.idx = 6
)
seg.plot[[1]]$segment_plot
```
\end{subfigure}
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:segplot2}scale\_ccf\_plot shows the cross-correlation curve between the reference signature and segment 6 at level one. The five highest peaks are marked by dots. The vertical red dashed line indicates the congruent registration position; the green dashed line shows a peak position in the highest segment level; the blue dashed lines show the tolerance zone around the green dashed line. We can see that none of the five highest peaks at level one falls within the tolerance zone, indicating that there is no consistent correlation peak or a candidate position identified by basis segment 6 under the multi-segment lengths strategy. Thus, the basis segment 6 doesn't vote for the congruent registration position and is not a cmps.}
```{r segplot2, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE}
seg.plot[[1]]$scale_ccf_plot
```
\end{subfigure}
\caption{\label{fig:segplots} Two plots used to investigate basis segment 6 at level one}
\end{figure}
```
Additionally, users can have more insights about why segment 6 is not a congruent matching profile segment if we put the `segment_plot` and `scale_ccf_plot` of all three segment levels together, as shown in \autoref{fig:seg_all} with the help of `ggpubr::ggarrange()`.

```{=tex}
\begin{figure}
```{r segplot3, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE}
library(ggpubr)

ggarrange(
  plotlist = 
    unlist(seg.plot, 
           recursive = FALSE),
  ncol = 2, 
  nrow = 3)
```
\caption{\label{fig:seg_all} Put segment\_plot and scale\_ccf\_plot of all three levels together. We are identifying five highest peaks at level one, three peaks at level two, and one peak at level three since npeaks.set = c(5, 3, 1). The highest peak position at level three is marked by the green dashed line across all segment levels. However, the highest peak on level three does not coincide with any of the top five highest peaks at level one. This indicates that there is no consistent correlation peak or a candidate position for basis segment 6 under the multi-segment lengths strategy.}
\end{figure}
```
In \autoref{fig:seg_all}, the red vertical dashed line indicates the congruent registration position.
We can see that the basis segment 6 does obtain a peak near the congruent registration position at level two and level three, respectively; however, this position doesn't give one of the five highest peaks at level one.
As a result, segment 6 fails to identify the consistent correlation peak (ccp) and fails to become one of the congruent matching profile segments according to the multi-segment lengths strategy.
The identified top five peaks at level one are also examples of "false positive" peaks.
The "true positive" peak (the peak within the tolerance zone of the congruent registration position) is identified at level two and three by increasing the segment length, which justifies the usage of multi-segment lengths strategy.

## Evaluation Metrics

**Metrics Based on CMPS scores**

>>>>

---
title: "cover_letter"
author: "xxx"
date: "10/13/2021"
output: html_document
---
